\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-03-21 paper.tex
update ISTA

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{Proximal gradient descent ISTA$_{\text{sg} l_1}$($v$, $f$, $\tilde \eta$, $\zeta$)}
\label{alg:ISTA}
\KwInput{A vector $v$, a function $f$ (its gradient), hyper-parameters $\tilde \eta$ and $\zeta$}
\KwOutput{$\flat^{\breve{w}^{max}}$}
Initialize $\flat^{(0)}=v$ \\
\For{$\breve{w} = 0,\ldots,\breve{w}^{max}$}    
{ 
  $\begin{aligned}
  \flat^{(\breve{w}+1)} \in &\argmin_u \Big\{\dfrac{1}{2\varrho_{\breve{w}}} \left\lVert u - \big(\flat^{(\breve{w})} - \varrho_{\breve{w}} \dfrac{\partial f(u)}{\partial u} \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\} \\
  &\coloneqq\text{prox}_{\text{sg} l_1, \tilde{\eta}, \zeta} \Big(\flat^{(\breve{w})} - \varrho_{\breve{w}} \dfrac{\partial f(u)}{\partial u} \Big)
  \end{aligned}$
}
\textbf{Return:} {$\flat^{\breve{w}^{max}}$}
\end{algorithm}

\vspace{.5cm}
TODO w/ VT: note on $\breve{w}^{max}$ and $\varrho_{\breve{w}}$. Lemma~\ref{lemma:prox-sgl1} bellow states that one can easily compute the proximal operator of the sparse group lasso defined in Algorithm~\ref{alg:ISTA}.
\begin{lemma}
\label{lemma:prox-sgl1}
The proximal operator of the sparse group lasso can be expressed as the composition of the group lasso and the lasso proximal operators, that is
\normalfont
\begin{equation*}
  \text{prox}_{\text{sg} l_1,\tilde \eta, \zeta} = \text{prox}_{\zeta \tilde \eta \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\zeta (1 - \tilde \eta) \norm{\cdot}_1}.
\end{equation*}
\end{lemma}
The two proximal operators on the right hand side of Lemma~\ref{lemma:prox-sgl1} are both well known analytically~\citep{bach2011optimization} and tractable. We used the \texttt{Python} library \texttt{copt} for the implementation of proximal gradient descent~\citep{copt} and we propose in our package an implementation for the proximal operator of the sparse group lasso based on Lemma~\ref{lemma:prox-sgl1}.
A proof of Lemma~\ref{lemma:prox-sgl1} is proposed in Appendix~\ref{sec:proof-lemma-prox-sgl1}.
Hence, we update $\beta_k^{(w)}$ with the following step
\begin{equation}
  \label{eq:beta-update}
  \beta_k^{(w+1)} = \text{ISTA}_{\text{sg} l_1}(\beta_k^{(w)}, R^{(w)}_{n,k}, \tilde \eta, \zeta_{2,k}).
\end{equation}
Note that the warmstart initialization is left to the user's choice, as well as the ISTA acceleration~\citep{beck2009fast}, since warmstart can diminish acceleration effectiveness~\citep{tibshirani2010proximal}.
The gradient of $R_{n,k}^{(w)}$ for the Gaussian case is here given by
\begin{align}
  \label{eq:grad-beta_k}
  \frac{\partial R_{n,k}^{(w)}(\beta_k)}{\partial \beta_k^l} = -n^{-1} &\sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[\sum_{a=1}^\mathcal{A} {\gamma_{k,a}^l}^{(w)} \Big( \delta_i  \hat \E_{\theta^{(w)}}[ \frac{\partial \varphi_a(t_i, \beta_k^l, b_i^l)}{\partial \beta_k^l} | t_i, \delta_i, y_i] \nonumber \\ 
  &- \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \frac{\partial \varphi_a(t_j, \beta_k^l, b_i^l)}{\partial \beta_k^l} \tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k^{(w)}, \beta_k}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big) \nonumber \\
  &+ U_{il}^\top {\phi_l^{(w)}}^{-1}I_{n_{il}} (y_i^l - U_{il}\beta_k^l - V_{il}\hat \E_{\theta^{(w)}}[ b_i | t_i, \delta_i, y_i])\Big]
\end{align}
for all $l = 1, \ldots, L$. Note that the gradients of the considered $\varphi_a$ functions with respect to $\beta_k^l$ are given in Table~\ref{table:shared_associations}.

>>>>

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{Proximal gradient descent ISTA$_{\text{sg} l_1}$($v$, $f$, $\tilde \eta$, $\zeta$)}
\label{alg:ISTA}
\KwInput{A vector $v$, a function $f$ (its gradient $\nabla f$), hyper-parameters $\tilde \eta$, $tol$ and $\zeta$}
\KwOutput{$\flat^{\breve{w}^{max}}$}
Define $p_{\varrho}(y) = \argmin_u \Big\{\dfrac{1}{2\varrho} \left\lVert u - \big(y - \varrho  \nabla f(y) \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\}$\\
Initialize $\flat^{(0)}=v$, $\varrho^0 > 0$, some $\epsilon > 1$ \\
\For{$\breve{w} = 0,\ldots,$ until convergence}    
{Backtracking line search : find the smallest nonnegative integer $i^{\breve{w}}$ such that
$f\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})})\big) \leq f(\flat^{(\breve{w})}) + \big<\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}\big), \nabla f(\flat^{(\breve{w})})\big> + \dfrac{1}{2\tilde{\varrho}_{\breve{w}}}\norm{p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}}_2^2$ 
with $\tilde{\varrho}_{\breve{w}} = \epsilon^{i^{\breve{w}}}\varrho^{\breve{w}}$\\
Set $\varrho^{(\breve{w} + 1)} = \tilde{\varrho}_{\breve{w}}$ \\
$\flat^{(\breve{w}+1)} = p_{\varrho^{(\breve{w} + 1)}}(\flat^{(\breve{w})}) \coloneqq\text{prox}_{\text{sg} l_1, \tilde{\eta}, \zeta} \big(\flat^{(\breve{w})} - \varrho^{(\breve{w} + 1)} \nabla f(\flat^{(\breve{w})}) \big)$ \\
\If{$\dfrac{\norm{\flat^{(\breve{w}+1)} - \flat^{(\breve{w})}}_2}{\varrho^{(\breve{w} + 1)}} < tol$}
	{
    \textbf{Return:} {$\flat^{\breve{w}}$}
	}
}
\end{algorithm}

\vspace{.5cm}
Lemma~\ref{lemma:prox-sgl1} bellow states that one can easily compute the proximal operator of the sparse group lasso defined in Algorithm~\ref{alg:ISTA}.
\begin{lemma}
\label{lemma:prox-sgl1}
The proximal operator of the sparse group lasso can be expressed as the composition of the group lasso and the lasso proximal operators, that is
\normalfont
\begin{equation*}
  \text{prox}_{\text{sg} l_1,\tilde \eta, \zeta} = \text{prox}_{\zeta \tilde \eta \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\zeta (1 - \tilde \eta) \norm{\cdot}_1}.
\end{equation*}
\end{lemma}
The two proximal operators on the right hand side of Lemma~\ref{lemma:prox-sgl1} are both well known analytically~\citep{bach2011optimization} and tractable. We used the \texttt{Python} library \texttt{copt} for the implementation of proximal gradient descent~\citep{copt} and we propose in our package an implementation for the proximal operator of the sparse group lasso based on Lemma~\ref{lemma:prox-sgl1}.
A proof of Lemma~\ref{lemma:prox-sgl1} is proposed in Appendix~\ref{sec:proof-lemma-prox-sgl1}.
Hence, we update $\beta_k^{(w)}$ with the following step
\begin{equation}
  \label{eq:beta-update}
  \beta_k^{(w+1)} = \text{ISTA}_{\text{sg} l_1}(\beta_k^{(w)}, R^{(w)}_{n,k}, \tilde \eta, \zeta_{2,k}).
\end{equation}
Note that the warmstart initialization is left to the user's choice, as well as the ISTA acceleration~\citep{beck2009fast}, since warmstart can diminish acceleration effectiveness~\citep{tibshirani2010proximal}.
The gradient of $R_{n,k}^{(w)}$ for the Gaussian case is here given by
\begin{align}
  \label{eq:grad-beta_k}
  \frac{\partial R_{n,k}^{(w)}(\beta_k)}{\partial \beta_k^l} = -n^{-1} &\sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[\sum_{a=1}^\mathcal{A} {\gamma_{k,a}^l}^{(w)} \Big( \delta_i  \hat \E_{\theta^{(w)}}[ \frac{\partial \varphi_a(t_i, \beta_k^l, b_i^l)}{\partial \beta_k^l} | t_i, \delta_i, y_i] \nonumber \\ 
  &- \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \frac{\partial \varphi_a(t_j, \beta_k^l, b_i^l)}{\partial \beta_k^l} \tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k^{(w)}, \beta_k}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big) \nonumber \\
  &+ U_{il}^\top {\phi_l^{(w)}}^{-1}I_{n_{il}} (y_i^l - U_{il}\beta_k^l - V_{il}\hat \E_{\theta^{(w)}}[ b_i | t_i, \delta_i, y_i])\Big]
\end{align}
for all $l = 1, \ldots, L$. Note that the gradients of the considered $\varphi_a$ functions with respect to $\beta_k^l$ are given in Table~\ref{table:shared_associations}.
